# Group table rows with panel  {#panel-intro}

## Syntax

```{r, eval = FALSE}
stable(stdata(), panel = "STUDY") 
```

To set a prefix for the panel header:
```{r, eval = FALSE}
stable(stdata(), panel = as.panel("STUDY", prefix = "Study: "))
```


## Basics

```{r, include = FALSE}
knitr::opts_chunk$set(comment = '.')
source("global.R")
```

```{r, include = FALSE}
smcars <- mutate(mtcars, name = rownames(mtcars))
smcars <- select(smcars, name, everything())
smcars <- filter(smcars, cyl < 8)
smcars <- arrange(smcars, carb)
smcars <- slice(smcars, 1:10)
```

Paneling your table is a way to group sets of rows together into a "panel" 
with a panel header in bold. For example, we can panel a table of `mtcars` by
`carb`.  We will be working with an abbreviated version of `mtcars`:

```{r}
smcars
```

Then we pass into `stable()` and name the paneling column:
```{r}
smcars %>% stable(panel = "carb") %>% st_asis()
```

Now, all of the `carb==1` rows are grouped with the heading `1` and similarly 
the `carb==2` rows are grouped with the heading `2` in bold. 

This is ok, but a more informative heading would be helpful. To do this, we'll 
call `as.panel()` to both name the panel column and set some options:

```{r}
```{r}
smcars %>% stable(panel = as.panel("carb", prefix = "carb: ")) %>% st_asis()
```
`````





## panel: important points

1. Most of the time, the data frame should be sorted by the `panel` column
1. 


\clearpage
