# pmtable

__Some setup__

```{r}
units = ys_get_unit(ys_help$spec(), parens = TRUE)
```

```{r}
data <- pmt_first
data_pk <- pmt_pk
data_all <- pmt_obs
```


## Data inventory tables

- Count number of
  - individuals
  - observations
  - BQL observations
  - missing values
- Calculate the percent  of observations or BQL in different sub groups


\clearpage

### Stacked by endpoint

- The stacked plot creates multiple independent tables to summarize different
endpoints; there is no single overall summary for the table because we
are summarizing different endpoints



```{r}
out <- pt_data_inventory(
  data_all,
  by = c(Study = "STUDYf"),
  panel = as.panel("SEQf", prefix = "Endpoint: "),
  stacked = TRUE
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage

### Paneled

- Just summarize a single endpoint

```{r}
out <- pt_data_inventory(
  data_pk,
  by = c(Study = "STUDYf"),
  panel = "ASIANf"
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage

### Grouped (by study)

```{r}
out <- pt_data_inventory(
  data_pk,
  by = c(Study = "STUDYf")
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage


## Wide categorical table

- Summary of categorical data in wide format
- The summary is `number (percent within group)`
- Wide refers to the fact that the covariates go across the table


#### Ungrouped

```{r}
out <- pt_cat_wide(
  data = data,
  cols = vars(Formulation = FORMf, Sex = SEXf, "Race group" = ASIANf)
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage

### Paneled (limited utility, IMO)

- Provided here for completeness

```{r}
out <- pt_cat_wide(
  data = data,
  cols = vars(Formulation = FORMf, Sex = SEXf, "Race group" = ASIANf),
  panel = as.panel("STUDYf", prefix = "Study: ")
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```



\clearpage

### Grouped (by male / female)

```{r}
out <- pt_cat_wide(
  data = data,
  by = c(Sex = "SEXf"),
  cols = vars(Formulation = FORMf, "Race group" = ASIANf)
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage

### Paneled and grouped

```{r}
out <- pt_cat_wide(
  data = data,
  cols = vars(Formulation = FORMf, Sex = SEXf,"Race group" = ASIANf),
  panel = as.panel("STUDYf", prefix = "Study: "),
  by = c("RF Group" = "RFf")
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage

### No summary

```{r}
out <- pt_cat_wide(
  data = data,
  summarize = "none",
  cols = vars(Formulation = FORMf, Sex = SEXf,"Race group" = ASIANf),
  panel = as.panel("STUDYf", prefix = "Study: "),
  by = c("RF Group" = "RFf")
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage


## Long categorical table

- Categorical table in long format
- Long indicates that the covariates go down the table

### Ungrouped
```{r}
out <- pt_cat_long(
  data = data,
  cols = vars(Study = STUDYf, Sex = SEXf, "Race group" = ASIANf, "Child-Pugh" = CPf)
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
````

\clearpage

### Grouped (by formulation)

```{r}
out <- pt_cat_long(
  data = data,
  cols = vars(Study = STUDYf,Sex = SEXf,"Race group" = ASIANf, "Child-Pugh" = CPf),
  span = c(Formulation = "FORMf")
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage

### Summary on bottom and right

```{r}
out <- pt_cat_long(
  data = data,
  summarize = "both",
  cols = vars(Formulation = FORMf, Sex = SEXf,"Race group" = ASIANf),
  span = vars(Study = STUDYf)
  )

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage

### No summary

```{r}
out <- pt_cat_long(
  data = data,
  summarize = "none",
  cols = vars(Formulation = FORMf, Sex = SEXf,"Race group" = ASIANf),
  span = vars(Study = STUDYf)
  )

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage



## Wide continuous table

- Continuous table in wide format
- Wide means that the covariates go across the table


### Ungrouped

```{r}
out <- pt_cont_wide(
  data = data,
  cols = "WT,SCR,AGE,ALB,HT",
  units = units
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

### Paneled

```{r}
out <- pt_cont_wide(
  data = data,
  cols = "WT,SCR,AGE,ALB,HT",
  panel = c(Study = "STUDYf"),
  units = units
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage

### Grouped (by study)


```{r}
out <- pt_cont_wide(
  data = data,
  cols = "WT,SCR,AGE,ALB,HT",
  by = c(Study = "STUDYf"),
  units = units
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage

### Paneled and grouped

```{r}
out <- pt_cont_wide(
  data = data,
  cols = "WT,SCR,AGE,ALB,HT",
  by = c(Study = "STUDYf"),
  panel = c(Formulation = "FORMf"),
  units = units
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage


## Long continuous table

- Continuous summary table in long format
- Long indicates that covariates go down the table


### Ungrouped

```{r}
out <- pt_cont_long(
  data = data,
  cols = "WT,SCR,AGE",
  units = units
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```

\clearpage

### Paneled

```{r}
out <- pt_cont_long(
  data = data,
  cols = "WT,SCR,AGE",
  panel = vars(Study = STUDYf),
  units = units
)

out %>% stable(r_file = "test.R", output_file = "test.tex") %>% st_asis()
```
