
# Options for previewing the table

## st2report

Take a table or a list of tables and render them in a report-like document with 
a table of contents and a (faux) caption for each table

```{r, eval = FALSE}
data %>% stable() %>% st2report()
```

You might try passing `ntex` to force the document to build more than once
(sometimes the layout settles down after the second build
```{r, eval = FALSE}
data %>% stable() %>% st2report(ntex = 2)
```

Pass in a list of tables and you will get one table on each page with a listing
of tables in the table of contents
```{r, eval = FALSE}
list(table1, table2, table3) %>% st2report()
```

## st2viewer

This function relies on `texPreview()` to render your table and display it as 
a graphic in the viewer window

```{r, eval = FALSE}
data %>% stable() %>% st2viewer()
```

This method is more conveient because the tables always go to the viewer. But 
the rendering will not be like what you will see in the report. 

## st2article

Like `st2report()` but less report-like. You should use `st2report()` instead.

```{r, eval = FALSE}
data %>% stable() %>% st2article()
```


## st2doc

The original.  Rather than building a TeX article, it runs the table in a 
`Rmd` document via pandoc.  Not recommended; it is much slower to get the 
preview because there has to be a call to `pandoc`

```{r, eval = FALSE}
data %>% stable() %>% st2doc()
```



