# Tables that span multiple pages: longtable

## Syntax

To create a long table from a data frame

```{r, eval = FALSE}
stable_long(stdata())
```

To create a long table from pipeline

```{r, eval = FALSE}
st_new(data) %>% stable_long()
```

To create a long table from `pmtable`

```{r, eval = FALSE}
pt_cont_long(data, cols = "WT,EGFR") %>% stable_long()
```


## Basics

You can create `longtables` that span multiple pages of your pdf document.
Tables using the`longtable` environment are very different than the the 
basic table from `stable()` which are built using tabular environment.


## Inserting longtable into your latex document

Once you have written your long table out to a file, you can source it into 
your latex document with a simple input command

```{txt, eval = FALSE}
\input{my-table.tex}
```

*IMPORTANT*  

- Do not wrap the input in `\begin{table}` / `\end{table}`; the table
will not show up properly that way
- Do not include any `\caption{...}` statement; the caption must be
provided in a special way (see below)

## Writing a caption

Longtable is different than regular tabular table in that the caption 
(and label) need to be included in the longtable environment.  This means
that you have to enter this information **when you create the table**.


### Method 1: in the R script

`stable_long()` has an argument called `lt_cap_text` that will allow you
to provide the text for the caption. This text must be passed to the
`stable_long()` call (or `st_make()`)

```{r, eval = FALSE}
out <- stable_long(data, lt_cap_text = "A long table (example 3)")
```

You will see in the TeX code that a caption is included in the longtable
environemt.


### Method 2: as a TeX macro

Rather than passing the text for the caption, you can name a macro that
should (will) be defined at the time the table is rendered in the TeX
document.

Use the `lt_cap_macro` argument:

```{r, eval = FALSE}
out <- stable_long(data, lt_cap_macro = "ltexfourcap")
```

Now, there is a call to that macro in the table and you **must** define that 
macro prior to sourcing in your tex document.

```{txt}
\newcommand{\ltexfourcap}{
  Another long table - example 4  
}

\input{example4.tex}
```

## Add a label

To add a label, pass in a caption (either as text or as a macro) and
then use the `lt_cap_label` argument:

```{r, eval = FALSE}
data %>% 
  stable_long(lt_cap_label = "tab:example", lt_cap_macro = "example") 
```

## Include a longtable in an Rmd document

If you want to dump a table into an R markdown document, pass it to
`st_asis()`. There is a method for `stable_long()` and it will wrap the
table properly.

```{r, eval = FALSE}
out <- stable_long(data) %>% st_asis()
```

