---
title: Preview
subtitle: Options for previewing your table.
toc-depth: 3
---


## st2report


Take a table or a list of tables and render them in a report-like document with 
a table of contents and a (faux) caption for each table

```{r, eval = FALSE}
data %>% stable() %>% st2report()
```

You might try passing `ntex` to force the document to build more than once
(sometimes the layout settles down after the second build

```{r, eval = FALSE}
data %>% stable() %>% st2report(ntex = 2)
```

## Multiple tables 


Pass in a list of tables and you will get one table on each page with a listing
of tables in the table of contents

```{r, eval = FALSE}
list(table1, table2, table3) %>% st2report()
```

When you pass a named list, those names will propagate into the preview 
table of contents. For example

```{r, eval = FALSE}
tab <- stable(stdata())

tablist <- list("first table" = tab, "second table" = tab)

tablist %>% st2report(ntex = 2)
```

You will most likely need to run `pdflatex` x2 to get the table of contents 
right.


## Landscape

Pass the `stable()` output through `as_lscape()` to have the preview page
render in landscape orientation

```{r, eval = FALSE}
stdata() %>% stable() %>% as_lscape() %>% st2report()
```

Note that this landscaping only operates in the preview; if you want your 
table to render in landscape mode in a standalone report, you will have to
code that in the report document.

## Preview other


### st2viewer

This function relies on `texPreview()` to render your table and display it as 
a graphic in the viewer window

```{r, eval = FALSE}
data %>% stable() %>% st2viewer()
```

This method is more convenient because the tables always go to the viewer. But 
the rendering will not be like what you will see in the report. 

### st2article

Like `st2report()` but less report-like. You should use `st2report()` instead.

```{r, eval = FALSE}
data %>% stable() %>% st2article()
```


### st2doc

The original.  Rather than building a TeX article, it runs the table in a 
`Rmd` document via pandoc.  Not recommended; it is much slower to get the 
preview because there has to be a call to `pandoc`

```{r, eval = FALSE}
data %>% stable() %>% st2doc()
```
